---
title: "Plot CTA1 comp ASR"
output: html_document
date: "2023-08-05"
---
# Import data
```{r}
library(tidyverse)
raw <- read.table("/space/htang5/Downloads/Thesis/Chapter 2-ASR Species/Input/CaASR-CH2.tsv", sep="\t", header=TRUE)
 print(raw)
```

```{r}
library(dplyr)
tmp <-  raw %>%
  mutate(
    scaled = Count * Dilutions * 1e-2
  ) %>% 
  # remove uninformative columns. only one H2O2 conc used for each species
  select(-Len_1, -Len_2)
dat.f3 <- tmp %>% 
  # group by primary to calculate r (MO/MM) or r' (PO/PM)
  separate(Group, into = c("Primary", "Secondary"), sep = 1) %>% 
  group_by(Date, Strain, Primary) %>% 
  # calculate % survival
  mutate(scaled_M = scaled[Secondary == "M"],
         r = num(scaled / scaled[Secondary == "M"], digits = 3))%>% 
   #filter(Secondary != "M") %>%
  #pivot_wider(id_cols = c(Date, Strain, Species, H2O2), 
              #names_from = Primary, values_from = r, names_prefix = "r") %>% 
  #mutate(ASR_score = rP / rM) %>%
mutate(Label = paste(Strain, H2O2, sep = "_")) %>%
print(dat.f3)
#write_tsv(dat.f3, file = "/space/htang5/Downloads/CaASR-Ch2-Calculated.tsv")
```


### Plot (all data)
```{r}
library(ggplot2)
require(tidyverse)
require(cowplot)
require(broom)

dat.f3 %>% 
  mutate( Primary = factor(Primary, levels = c("M", "P"), 
                     labels = c("Mock", "-Pi")),
        Species = factor(Species, levels = c("Ca", "Cg", "Sc"),
                      labels = c("C. albicans", "C. glabrata", "S. cerevisiae")),
        H2O2 = paste(H2O2, " mM")
  ) %>%
  filter(Label %in% c("yH714_6", "yH154_4")) %>%
  ggplot(aes(x = Primary, y = r)) + 
  geom_point(aes(shape = Primary), stroke = 1, size = 2, 
             position = position_dodge(0.9)) +
  stat_summary(aes(group = Primary), position = position_dodge(0.9),
               fun = mean, fun.max = mean, fun.min = mean,
               geom = "crossbar", color = "red", width = 0.5) +
  facet_wrap(~ Species + H2O2, nrow = 1, scales = "free_x") +
  scale_shape_manual(values = c("Mock" = 1, "-Pi" = 16)) +
  scale_y_continuous(limits = c(0, 0.2), labels = scales::percent) +
  xlab("Primary stress (45 min)") + ylab("% survival") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        axis.title.x = element_blank(), 
        strip.background = element_blank(),
        axis.text.x = element_text(size = 12, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 12, face = "bold", color = "black"),
        axis.title.y = element_text(size = 12, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 14, face = "bold")) 
ggsave("/space/htang5/Downloads/Thesis/Chapter 2-ASR Species/Output/20240122-CaASR-Plot.png")
```
>Plot Basal Survival

```{r}
dat.f3 %>% 
  dplyr::filter(Species %in% c("Ca", "Sc")) %>% 
  dplyr::filter(H2O2 %in% c("2", "4", "6"))  %>%
  mutate( Primary = factor(Primary, levels = c("M", "P"), 
                     labels = c("Mock", "-Pi")),
        H2O2 = paste(H2O2, " mM")
  ) %>%
  ggplot(aes(x = Primary, y = r)) + 
  geom_point(aes(shape = Primary), stroke = 1, size = 2, 
             position = position_dodge(0.9)) +
  stat_summary(aes(group = Primary), position = position_dodge(0.9),
               fun = mean, fun.max = mean, fun.min = mean,
               geom = "crossbar", color = "red", width = 0.5) +
  facet_wrap(~ Species + H2O2, nrow = 1, scales = "free_x") +
  scale_shape_manual(values = c("Mock" = 1, "-Pi" = 16)) +
  scale_y_continuous(limits = c(0, 0.2), labels = scales::percent) +
  xlab("Primary stress (45 min)") + ylab("% survival") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        axis.title.x = element_blank(), 
        strip.background = element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_text(size = 12, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 12, face = "bold", color = "black"),
        axis.title.y = element_text(size = 12, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 14, face = "bold")) 
ggsave("/space/htang5/Downloads/Thesis/Chapter 2-ASR Species/Output/20240517-CaASR-Basal-Plot.png")
```



# Statistical Tests

> The basal survival rates between species within the same day are not "paired". We will use a rank-sum test here.

```{r}
tmp <- dat.f3 %>% 
  dplyr::filter(Label %in% c("yH714_6", "yH154_4")) %>% 
  dplyr::filter(Primary == "M") %>% 
  pivot_wider(id_cols = c(Date), names_from = Species, values_from = r)
tmp
with(tmp, t.test(as.numeric(Ca), as.numeric(Sc), paired = FALSE))

with(tmp, wilcox.test(Ca, Sc, paired = FALSE))
```
> p = 0.73, basal survival rates are not significantly between species

_Primary stress enhanced survival (ASR) in Ca_

The comparison between r and r' is paired. We will use a signed-rank test.
```{r}
tmp <- dat.f3 %>% 
  dplyr::filter(Label %in% c("yH714_6")) %>% 
  pivot_wider(id_cols = c(Date), names_from = Primary, values_from = r) %>% 
  mutate(ASR = P/M)
tmp

x <- Hmisc::smean.cl.boot(tmp$ASR)
sprintf("ASR_score mean = %.2f, 95%% CI by bootstrap = [%.2f, %.2f]", x[1], x[2], x[3])

with(tmp, t.test(as.numeric(P), as.numeric(M), paired = TRUE, alternative = "g"))

with(tmp, wilcox.test(P, M, paired = TRUE, alternative = "g"))
```
> p = 0.0625, P and M difference is not significant in Ca

_Primary stress no effect in Sc_

```{r}
tmp <- dat.f3 %>% 
  dplyr::filter(Label %in% c("yH154_4")) %>% 
  pivot_wider(id_cols = c(Date), names_from = Primary, values_from = r) %>% 
  mutate(ASR = P/M)
tmp

x <- Hmisc::smean.cl.boot(tmp$ASR)

sprintf("ASR_score mean = %.2f, 95%% CI by bootstrap = [%.2f, %.2f]", x[1], x[2], x[3])

with(tmp, t.test(as.numeric(P), as.numeric(M), paired = TRUE, alternative = "g"))

with(tmp, wilcox.test(P, M, paired = TRUE, alternative = "g"))
```
> p = 0.087, P and M difference is not significant in Sc




