---
title: "SYTO9-Cell-Density"
output: html_document
date: "2023-12-21"
---
# Background
This is the flow cytometry mock treated Cg stained with 1000x dilution of SYTO9 in 0.01-1OD cell density and run through flow (details in HT's ELN). Goal is to show SYTO9's fluorescence intensity change with cell density.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
data.path = "/space/htang5/Documents/122123_Cell Density/Raw"
dat.f1d <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
require(PTXQC)
oriNames <- sampleNames(dat.f1d)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 6, 7, 8)] 
colnames(tmp) <- c("Date", "Treatment", "Density", "Dye") 
sample <- data.frame(tmp) %>% 
  mutate(Dye = gsub(".fcs", "", Dye) %>% as.integer(),
         Treatment = factor(Treatment, levels = (c("0mM", "10mM", "100mM", "1M"))))
rownames(sample) <- oriNames
pData(dat.f1d) <- sample
print(pData(dat.f1d))
write_tsv(pData(dat.f1d), file = "../Input/20240104-fig-1d-cell-density-raw-data-ht.tsv")
```

# Extract MFI across flowFrames
```{r}
tmp <- fsApply(dat.f1d, each_col, median)
if(all(rownames(tmp) == rownames(sample))){
  combined_data <- cbind(pData(dat.f1d), tmp) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
}
write_tsv(combined_data, file = "../Input/20240104-fig-1d-MFI-ht.tsv")
print(combined_data)
```

# Graph Green MFIs against teatment and group by dye concentration

```{r}
pf1d <- combined_data %>%
  dplyr::filter(Density != "U") %>% 
  dplyr::filter(Treatment == "0mM") %>%
  ggplot(aes(x = Density, y = BL1.H)) +
  facet_wrap(~ Dye, scales = "free_y", labeller = as_labeller(c("500" = "500x", "1000" = "1000x", "2000" = "2000x"))) +
  geom_boxplot() +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.2)) +
  theme_minimal()
pf1d +
  labs(x = "Cell Density (OD)", y = "Median Green Fluorescence Intensity") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 16, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 16, face = "bold", color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 16, face = "bold")) 
ggsave("../output/20240104-SYTO9-Cell-Density-Green-MFI-Plot.png", width = 7, height = 5)
```
# Graph Red MFIs against teatment and group by dye concentration

```{r}
pf1d <- combined_data %>%
  dplyr::filter(Density != "U") %>% 
  dplyr::filter(Treatment == "0mM") %>%
  ggplot(aes(x = Density, y = BL3.H)) +
  facet_wrap(~ Dye, scales = "free_y", labeller = as_labeller(c("500" = "500x", "1000" = "1000x", "2000" = "2000x"))) +
  geom_boxplot() +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.2)) +
  theme_minimal()
pf1d +
  labs(x = "Cell Density (OD)", y = "Median Red Fluorescence Intensity") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 16, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 16, face = "bold", color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 16, face = "bold")) 
ggsave("../output/20240104-SYTO9-Cell-Density-Red-MFI-Plot.png", width = 7, height = 5)
```

