---
title: "Buffer Optimization"
output: html_document
date: "2023-12-18"
---

# Background
This is the flow cytometry data for miglog phase Cg stained with PI/SYTO9/PI+SYTO9 in different staining buffers, and run through flow cytometry (details in ELN). The goal is to establish FungaLight's staining pattern in various staining buffers.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
require(ggrdiges)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "/space/htang5/Documents/0108-1024_Extended buffer optimization"
dat.f1e <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
require(PTXQC)
oriNames <- sampleNames(dat.f1e)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 6, 7, 8)] 
colnames(tmp) <- c("Date", "Buffer", "Treatment", "Dye") 
sample <- data.frame(tmp) %>% 
  mutate(
    Dye = ifelse(Dye == "p.fcs", "PI", ifelse(Dye == "s.fcs", "SYTO9", "Both")),
    Buffer = ifelse(Buffer == "sa", "Saline", ifelse(Buffer == "pb", "PBS", ifelse(Buffer == "w", "ddH2O", "sc complete"))),
    Treatment = factor(Treatment, levels = as.character(c(0, 10, 100, 1000))) %>% paste(" mM", sep = "")
  )
rownames(sample) <- oriNames
pData(dat.f1e) <- sample
print(pData(dat.f1e))
write_tsv(pData(dat.f1e), file = "../Input/20240111-fig-1-buffer-raw-data-ht.tsv")
```

# Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier 

```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat.f1e [1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat.f1e) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

>Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat.f1e[[1]], outlier.gate)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

```{r}
print(dat.f1e)
```


# Output Graph Production
>Several output graphs are produced, and are stored in the Output folder.

```{r}
sub <- dat.f1e[pData(dat.f1e)$Dye == "PI" & pData(dat.f1e)$Date == "010924"] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  facet_grid(Buffer ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM",
                                      "Saline" = "0.85% Saline", "PBS" = "PBS", "ddH2O" = "DI Water", "sc complete" = "sc Media"))) + 
  theme_minimal()
p2 + scale_x_logicle() + scale_y_logicle() +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence") +
  panel_border(color = "black", size = 1) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        ) + 
  theme(strip.text = element_text(size = 9, face = "bold")) 
ggsave("../output/20240111-PI-Buffers-Flow-Plot.png", width = 7, height = 5)
```

```{r}
sub <- dat.f1e[pData(dat.f1e)$Dye == "SYTO9" & pData(dat.f1e)$Date == "010824"] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  facet_grid(Buffer ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM",
                                      "Saline" = "0.85% Saline", "PBS" = "PBS", "ddH2O" = "DI Water", "sc complete" = "sc Media")))  + 
  theme_minimal()
p2 + scale_x_logicle() + scale_y_logicle() +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence") +
  panel_border(color = "black", size = 1) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        ) + 
  theme(strip.text = element_text(size = 9, face = "bold")) 
ggsave("../output/20240111-SYTO9-Buffers-Flow-Plot.png", width = 7, height = 5)
```
```{r}
sub <- dat.f1e[pData(dat.f1e)$Dye == "Both" & pData(dat.f1e)$Date == "010824"] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  facet_grid(Buffer ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM",
                                      "Saline" = "0.85% Saline", "PBS" = "PBS", "ddH2O" = "DI Water", "sc complete" = "sc Media")))  + 
  theme_minimal()
p2 + scale_x_logicle() + scale_y_logicle() +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence") +
  panel_border(color = "black", size = 1) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        ) + 
  theme(strip.text = element_text(size = 9, face = "bold")) 
ggsave("../output/20240111-Both-Buffers-Flow-Plot.png", width = 7, height = 5)
```














