---
title: "PI-Timecourse-Graphing"
author: HT and BH
date: "2023-12-20 (updated `r Sys.Date()`)"
output:
  html_notebook:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: hide
---

# Background
This is the flow cytometry mock treated Cg stained with 1000x dilution of SYTO9 for 0-60min and run through flow (details in HT's ELN). Goal is to show SYTO9's fluorescence intensity change with time.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
data.path = "/space/htang5/Documents/20240506-09"
dat.f1c <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify Sample Names
```{r}
oriNames <- sampleNames(dat.f1c)
tmp <- str_split(oriNames, pattern = "_", simplify = TRUE)[,c(1, 6)] 
colnames(tmp) <- c("Date", "Time") 
sample <- data.frame(tmp) %>%
  mutate(
    Time = gsub(".fcs", "", Time),
    Time = factor(Time, levels = as.character(c(0, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55)))
  )
rownames(sample) <- oriNames
pData(dat.f1c) <- sample
write_tsv(pData(dat.f1c), file = "../Input/20240508-PI-timecourse-data-ht.tsv")
```

# Extract MFI across flowFrames
>MFI are extracted and wrote into tsv. The tsv is stored in the input folder.

```{r}
tmp <- fsApply(dat.f1c, each_col, median)
if(all(rownames(tmp) == rownames(sample))){
  combined_data_f1c <- cbind(pData(dat.f1c), tmp) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
}
print(combined_data_f1c)
write_tsv(combined_data_f1c, file = "../Input/20240508-PI-timecourse-MFI-ht.tsv")
```

# Graphing 

> Graph medians against time

```{r}
combined_data_f1c %>%
  ggplot(aes(x = Time, y = BL3.H, group = 1)) +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.1)) +
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", size = 0.5, aes(group = Time), color = "red")+
  stat_summary(aes(group = Time), fun = mean, fun.max = mean, fun.min = mean, geom = "crossbar", color = "red", width = 0.5) +
  labs(x = "Time (min)", y = "Red MFI") +
  theme_minimal() +
  panel_border(color = "black", size = 1) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold")
        )
ggsave("../output/20240508-PI-Timecourse-Red-MFI-Plot.png", width = 10, height = 5)
```











