---
title: "FungaLight Pre-oprimization Figure"
output: html_document
date: "2023-12-18"
---

# Background
This is the flow cytometry data for miglog phase Cg stained with PI/SYTO9/PI+SYTO9 in different staining buffers, and run through flow cytometry (details in ELN). The goal is to establish FungaLight's staining pattern in various staining buffers.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
require(ggrdiges)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "/space/htang5/Documents/20231004"
dat.f1e <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
require(PTXQC)
oriNames <- sampleNames(dat.f1e)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 7)] 
colnames(tmp) <- c("Date", "Treatment") 
sample <- data.frame(tmp) %>% 
  mutate(
    Treatment = gsub("\\.fcs", "", Treatment),
    Treatment = gsub("1M", "1000", Treatment),
    Treatment = gsub("Mock", "0", Treatment),
    Treatment = factor(Treatment, levels = c("0", "2", "4", "6", "8", "10", "20", "40", "60", "80", "100", "1000")) %>%
      paste(" mM", sep = "")
  )
rownames(sample) <- oriNames
pData(dat.f1e) <- sample
print(pData(dat.f1e))
#write_tsv(pData(dat.f1e), file = "/space/htang5/Downloads/Thesis/Chapter 4 - FGL/Input/20240405-pre-opt-raw-data-ht.tsv")
```

# Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier 

```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat.f1e [1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat.f1e) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

>Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat.f1e[[1]], outlier.gate)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

# Gate for Unstained population
> Unstained population is defined as population below 10^2 in both channels (in the noise range).

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1e[[7]], singlet.gate)
polygon2 <- matrix(c(0, 1e2, 1e2, 0,
                    0, 0, 1e2, 1e2), ncol = 2)
colnames(polygon2) <- c("BL1.H", "BL3.H")
unstained.gate <- polygonGate(filterId = "unstained", boundaries = polygon2)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(unstained.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, unstained.gate, parent = "singlet", name = "unstained")
recompute(gs)
```

# Gate for Dead population
> Dead population is defined based on 1M treated sample.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1e[[10]], singlet.gate)
polygon <- matrix(c(10^3, 10^1.5,10^2.5, 10^4.5,
                    10^4.5, 10^2.2, 10^2.2, 10^4), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
dead.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(dead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, dead.gate, parent = "singlet", name = "dead")
recompute(gs)
```
# Gate for peak population

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1e[[21]], singlet.gate)
polygon2 <- matrix(c(10^3, 10^4.5, 10^3,
                    10^1.2, 10^3.5, 10^2.6), ncol = 2)
colnames(polygon2) <- c("BL1.H", "BL3.H")
peak.gate <- polygonGate(filterId = "peak", boundaries = polygon2)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(peak.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, peak.gate, parent = "singlet", name = "peak")
recompute(gs)
```

>Exported gated event counts

```{r}
gated_stats_f1g <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)
gated_stats_f1g$percent_Dead <- gated_stats_f1g$n_dead / gated_stats_f1g$n_singlet
gated_stats_f1g$percent_Unstained <- gated_stats_f1g$n_unstained / gated_stats_f1g$n_singlet
gated_data <- cbind(pData(dat.f1e), gated_stats_f1g) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
#write_tsv(gated_data, file = "../Input/20240325-Ca-PI-Both-Distinguishing-Power-gated-ht.tsv")
print(gated_data)
```

```{r}
print(dat.f1e)
```


# Output Graph Production
>Several output graphs are produced, and are stored in the Output folder.

```{r fig.width=9, fig.height=6}
sub <- dat.f1e[pData(dat.f1e)$Date %in% c("092123", "092523", "100423") &
  (pData(dat.f1e)$Treatment == "0 mM" |
                 pData(dat.f1e)$Treatment == "10 mM" )]
p2 <- sub %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  facet_grid(Treatment ~ Date , 
             labeller = as_labeller(c("0 mM" = "H2O2: 0 mM", "10 mM" = "H2O2: 10 mM", "100 mM" = "H2O2: 100 mM", "1000 mM" = "H2O2: 1000 mM", "092123" = "Replicate 1", "092523" = "Replicate 2", "100423" = "Replicate 3"))) + 
  theme_minimal()
p2 + scale_x_logicle(breaks = c(100, 1000, 10000)) + scale_y_logicle(breaks = c(100, 1000, 10000)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 2) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        panel.spacing = unit(1, "lines"), 
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("../output/20240424-Pre-opt-Flow-Plot.png")
```

```{r fig.width=15, fig.height=5}
sub <- dat.f1e[pData(dat.f1e)$Treatment == "0 mM"]
p2 <- sub %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  geom_gate(dead.gate, size = 4)  + 
  geom_stats(size = 12) +
  facet_grid(Treatment ~ Date , 
             labeller = as_labeller(c("0 mM" = "H2O2: 0 mM", "10 mM" = "H2O2: 10 mM", "100 mM" = "H2O2: 100 mM", "1000 mM" = "H2O2: 1000 mM", "092123" = "Replicate 1", "092523" = "Replicate 2", "100423" = "Replicate 3"))) + 
  theme_minimal()
p2 + scale_x_logicle(breaks = c(100, 1000, 10000)) + scale_y_logicle(breaks = c(100, 1000, 10000)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 2) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        panel.spacing = unit(1, "lines"), 
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("../output/20240424-Pre-opt-dead-Plot.png")
```

```{r fig.width=18, fig.height=12}
sub <- dat.f1e[pData(dat.f1e)$Treatment == "0 mM"|
                 pData(dat.f1e)$Treatment == "10 mM" ]
p2 <- sub %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  geom_gate(peak.gate, size = 4)  + 
  geom_stats(size = 12) +
  facet_grid(Treatment ~ Date , 
             labeller = as_labeller(c("0 mM" = "H2O2: 0 mM", "10 mM" = "H2O2: 10 mM", "100 mM" = "H2O2: 100 mM", "1000 mM" = "H2O2: 1000 mM", "092123" = "Replicate 1", "092523" = "Replicate 2", "100423" = "Replicate 3"))) + 
  theme_minimal()
p2 + scale_x_logicle() + scale_y_logicle() +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 20, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 20, face = "bold", color = "black"),
        axis.title.x = element_text(size = 20, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("../output/20240411-Pre-opt-peak-Plot.png")
```
```{r fig.width=18, fig.height=12}
sub <- dat.f1e[(pData(dat.f1e)$Treatment == "0 mM" |
                 pData(dat.f1e)$Treatment == "10 mM" |
                  pData(dat.f1e)$Treatment == "100 mM" |
                 pData(dat.f1e)$Treatment == "1000 mM")]
p2 <- sub %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  geom_gate(unstained.gate, size = 4)  + 
  geom_stats(adjust = c(80, 0.5), size = 12) +
  facet_grid(Date ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "H2O2: 0 mM", "10 mM" = "H2O2: 10 mM", "100 mM" = "H2O2: 100 mM", "1000 mM" = "H2O2: 1000 mM", "092123" = "Replicate 1", "092523" = "Replicate 2", "100423" = "Replicate 3"))) + 
  theme_minimal()
p2 + scale_x_logicle(breaks = c(100, 1000, 10000)) + scale_y_logicle(breaks = c(100, 1000, 10000)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 2) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        panel.spacing = unit(1, "lines"), 
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("../output/20240424-Pre-unstained-peak-Plot.png")
```


```{r fig.width=6, fig.height=6}
p1 <- gated_data %>% 
  dplyr::filter(Treatment == "0 mM") %>% 
  ggplot(aes(x = Date, y = percent_Dead))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "skyblue2", alpha = 0.7) +
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Replicates", y = "% Dead Population") +
  ggtitle("Dead") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot(font_size = 25) 
p1
ggsave("../output/20240523-dead-in-mock.png")
```

```{r fig.width=6, fig.height=6}
p2 <- gated_data %>% 
   dplyr::filter(Treatment %in% c("0 mM", "10 mM", "100 mM"))  %>%
  mutate(Treatment = gsub(" mM", "", Treatment)) %>%
  ggplot(aes(x = Treatment, y = percent_Unstained))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "skyblue2", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Unstained") +
  ggtitle("SYTO9 only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot(font_size = 25) 
  
p2
ggsave("../output/20240523-unstained.png")
```














