---
title: "SYTO9-Concentration-Graphing"
output: html_document
date: "2023-12-18"
---

# Background
This is the flow cytometry data for miglog phase Cg stained with various concentrations of SYTO9 alone, and run through flow cytometry (details in ELN). The goal is to establish SYTO9's staining pattern in midlog phase yeasts cells.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
require(ggrdiges)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "/space/htang5/Documents/112723_SP Dilutions/Raw"
dat.f1 <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
require(PTXQC)
oriNames <- sampleNames(dat.f1)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 5, 6, 7)] 
colnames(tmp) <- c("Date", "Treatment", "Dilution", "Dye") 
sample <- data.frame(tmp) %>% 
  mutate(Dye = ifelse(Dye == "p.fcs", "PI", "SYTO9"),
         Dilution = factor(gsub(".fcs", "", Dilution)) %>% paste(" x", sep = ""),
         Treatment = factor(Treatment, levels = as.character(c(0, 10, 100, 1000))) %>% paste(" mM", sep = "")) %>%
  mutate(Dilution = factor(Dilution, levels = c("250 x", "500 x", "1000 x", "2000 x")))
rownames(sample) <- oriNames
pData(dat.f1) <- sample
print(pData(dat.f1))
#write_tsv(pData(dat.f1), file = "../Input/20231228-fig-1-raw-data-ht.tsv")
```

# Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier 

```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat.f1 [1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat.f1) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

>Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat.f1[[1]], outlier.gate)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```
# Gate for Dead population
> Dead population is defined based on 1M treated sample.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1[[5]], singlet.gate)
polygon <- matrix(c(0, 10^2.5, 10^2.5, 0, 
                    10^2.2, 10^2.2, 10^5, 10^5), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
dead.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(dead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, dead.gate, parent = "singlet", name = "dead")
recompute(gs)
```

# Gate for green positive and red positive population

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1[[24]], singlet.gate)
polygon4 <- matrix(c(1e2, 1e6, 1e6, 1e2,
                    10^2, 10^2, 1e6, 1e6), ncol = 2)
colnames(polygon4) <- c("BL1.H", "BL3.H")
RP.gate <- polygonGate(filterId = "RP", boundaries = polygon4)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(RP.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, RP.gate, parent = "singlet", name = "Red_Positive")
recompute(gs)
```

Gate for live population

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1[[9]], singlet.gate)
polygon <- matrix(c(0, 10^5, 10^5, 0,
                    10^2.2, 10^2.2, 0, 0), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
live.gate <- polygonGate(filterId = "live", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(live.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, live.gate, parent = "singlet", name = "live")
recompute(gs)
```

```{r}
gated_stats_f1g <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)
gated_stats_f1g$percent_Live <- gated_stats_f1g$n_live / gated_stats_f1g$n_singlet
gated_stats_f1g$percent_RP <- gated_stats_f1g$n_Red_Positive / gated_stats_f1g$n_singlet
gated_data <- cbind(pData(dat.f1), gated_stats_f1g) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
write_tsv(gated_data, file = "../Input/20240430-Cg-SYTO9-gated-ht.tsv")
print(gated_data)
```

# Output Graph Production
>Several output graphs are produced, and are stored in the Output folder.

```{r}
sub <- dat.f1[pData(dat.f1)$Dye == "PI" & pData(dat.f1)$Date == "112023"] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
ggcyto(aes(x = BL3.H)) + geom_histogram() + scale_x_logicle() + scale_y_log10() +
  facet_grid(Dilution ~ Treatment) +
  labs(x = "BL3.H", y = "Count", title = "Flow Cytometry Data") +
  theme_minimal()
p2
```

```{r}
sub <- dat.f1[pData(dat.f1)$Dye == "PI" & pData(dat.f1)$Date == "112023"] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + geom_gate(dead.gate) + geom_stats(adjust = c(19, 0.1)) +
  facet_grid(Dilution ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM", 
                                      "250 x" = "250x", "500 x" = "500x", "1000 x" = "1000x", "2000 x" = "2000x"))) + 
  theme_minimal()
p2 + scale_x_logicle(breaks = c(100, 1000, 10000)) + scale_y_logicle(breaks = c(100, 1000, 10000)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 2) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
#ggsave("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/01-SYTO9-Midlog-Characterization/output/20240425-PI-Dilutions-Raw-Flow-Plot.png")
```
```{r fig.width=16, fig.height=4}
sub <- dat.f1[pData(dat.f1)$Dye == "PI" & pData(dat.f1)$Date == "112023" & pData(dat.f1)$Dilution == "1000 x"] 
sg <- Subset(sub, singlet.gate)
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + geom_gate(dead.gate, size = 1) + geom_stats(adjust = c(20, 0.1), size = 10) +
  facet_grid(Dilution ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM", 
                                      "250 x" = "250x", "500 x" = "500x", "1000 x" = "1000x", "2000 x" = "2000x"))) + 
  theme_minimal()
p2 + scale_x_logicle (breaks = c(100, 1000, 10000, 100000, 1000000)) + scale_y_logicle (breaks = c(100, 1000, 10000, 100000, 1000000)) +
  coord_cartesian(xlim = c(1, 1e6), ylim = c(1, 1e6)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 2) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/01-SYTO9-Midlog-Characterization/output/20240521-PI-1000x-Raw-Flow-Plot.png")
```

```{r fig.width=16, fig.height=4}
sub <- dat.f1[pData(dat.f1)$Dye == "SYTO9" & pData(dat.f1)$Date == "112023" & pData(dat.f1)$Dilution == "1000 x"] 
sg <- Subset(sub, singlet.gate)
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + geom_gate(RP.gate, size = 1) + geom_stats(adjust = c(0.05, 0.1), size = 10) +
  facet_grid(Dilution ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM", 
                                      "250 x" = "250x", "500 x" = "500x", "1000 x" = "1000x", "2000 x" = "2000x"))) + 
  theme_minimal()
p2 + scale_x_logicle (breaks = c(100, 1000, 10000, 100000, 1000000)) + scale_y_logicle (breaks = c(100, 1000, 10000, 100000, 1000000)) +
  coord_cartesian(xlim = c(1, 1e6), ylim = c(1, 1e6)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 2) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 20, face = "bold")) 
ggsave("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/01-SYTO9-Midlog-Characterization/output/20240521-SYTO9-1000x-Raw-Flow-Plot.png")
```

> PI quantifications

```{r}
gated_data %>% 
   dplyr::filter(Dye == "PI") %>% 
 mutate(Dilution = gsub("(\\d+) x", "\\1x", Dilution)) %>% 
  mutate(Dilution = factor(Dilution, levels = c("250x", "500x", "1000x", "2000x"))) %>% 
ggplot(aes(x = Dilution, y = percent_Dead)) +
  geom_point(size = 3, position = position_jitter(0.1)) + 
  scale_color_brewer(palette = "Dark2") +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red2",
               size = 0.8, position = position_nudge(x = 0.3)) +
  facet_wrap(~ Treatment, nrow = 1, scales = "free_x") +
  scale_y_continuous(labels = scales::percent) +
  xlab("PI Folds Dilution") + ylab("% Red Positive Gated")  +
    theme_minimal() +
  theme(strip.text = element_text(size = rel(1), face = 3))+
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 20, face = "bold", color = "black", angle = 90),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 20, face = "bold", color = "black"),
        axis.title.x = element_text(size = 20, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold")
        )+ theme(strip.text = element_text(size = 18, face = "bold")) 
ggsave("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/01-SYTO9-Midlog-Characterization/output/20240409-PI-Dilutions-Gated-Percentage-Plot.png", width = 10, height = 7)
```



```{r}
sub <- dat.f1[pData(dat.f1)$Dye == "SYTO9" & pData(dat.f1)$Date == "112023"]
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + geom_gate(RP1.gate) + geom_stats(adjust = c(0.01, 0.1)) +
  facet_grid(Dilution ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM", 
                                      "250 x" = "250x", "500 x" = "500x", "1000 x" = "1000x", "2000 x" = "2000x"))) + 
  theme_minimal()
p2 + scale_x_logicle() + scale_y_logicle() +
  coord_cartesian(xlim = c(1, 1e6), ylim = c(1, 1e6)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence") +
  panel_border(color = "black", size = 1) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        ) + 
  theme(strip.text = element_text(size = 12, face = "bold")) 
ggsave("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/01-SYTO9-Midlog-Characterization/output/20240409-SYTO9-Dilutions-Raw-Flow-Plot.png", width = 7, height = 5)
```

# Extract MFI across flowFrames
>MFI are extracted and wrote into tsv. The tsv is stored in the input folder.

```{r}
tmp <- fsApply(dat.f1, each_col, median)
if(all(rownames(tmp) == rownames(sample))){
  combined_data_f1 <- cbind(pData(dat.f1), tmp) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
}
#write_tsv(combined_data_f1, file = "../Input/20231228-fig-1-MFI-ht.tsv")
print(combined_data_f1)
```

# Graph Green MFIs against teatment and group by dye concentration
>Several output graphs are produced. The graphs are include in Figure 1 Panel A.

```{r}
combined_data_f1 %>%
  dplyr::filter(Treatment == "0 mM") %>%
   mutate(Label = paste(Dilution, Dye, sep = "_"),
         Signal = ifelse(Dye == "PI", BL3.H, BL1.H)) %>%
  mutate(Dilution = gsub("(\\d+) x", "\\1x", Dilution)) %>% 
  mutate(Dilution = factor(Dilution, levels = c("250x", "500x", "1000x", "2000x"))) %>% 
  ggplot(aes(x = Dilution, y = Signal, group = Dye, color = Dye))  + 
  geom_bar(aes(fill = Dye), stat = "summary", fun = "mean", position = position_dodge(), alpha = 0.7) +
  theme_minimal()  +
  xlab("Dye Folds Dilution") + ylab("Fluorescent Signal")  +
    theme_minimal() +
  panel_border(color = "black", size = 1.5) +
   theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 20, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 20, face = "bold", color = "black"),
        axis.title.x = element_text(size = 20, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold"),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 18, face = "bold"),  # Set legend text size and face
        strip.text = element_text(size = 18, face = "bold")  # Adjust size and face for strip text
  ) 
ggsave("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/01-SYTO9-Midlog-Characterization/output/20240411-SYTO9-vsPI-Plot.png")

```

```{r}
pf1a <- combined_data_f1 %>%
  dplyr::filter(Dye == "SYTO9") %>% 
  dplyr::filter(Dilution == "1000 x") %>%
  mutate(Treatment = gsub("1000 mM", "1 M", Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = c("0 mM", "10 mM", "100 mM", "1 M"))) %>%
  ggplot(aes(x = Treatment, y = BL1.H)) +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.2)) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red2",
               size = 0.8, position = position_nudge(x = 0.1)) +
  theme_minimal()
pf1a + labs(x = bquote(H[2]*O[2]), y = "Green MFI") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold")
        )
ggsave("../output/20240522-SYTO9-Dilutions-Green-MFI-Plot.png", width = 7, height = 5)
```

> Graph red MFI for various SYTO9 dilutions


```{r}
pf1a <- combined_data_f1 %>%
  dplyr::filter(Dye == "SYTO9") %>% 
  dplyr::filter(Dilution == "1000 x") %>%
  mutate(Treatment = gsub("1000 mM", "1 M", Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = c("0 mM", "10 mM", "100 mM", "1 M"))) %>%
  ggplot(aes(x = Treatment, y = BL3.H)) +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.2)) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red2",
               size = 0.8, position = position_nudge(x = 0.1)) +
  theme_minimal()
pf1a +
 labs(x = bquote(H[2]*O[2]), y = "Red MFI") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold")
        )
ggsave("../output/20240522-SYTO9-Dilutions-Red-MFI-Plot.png", width = 7, height = 5)
```

> Graph a scatter plot for median red and greens for each population
No output from this plotting.

```{r}
combined_data_f1 %>%
  dplyr::filter(Dye == "SYTO9") %>% 
  dplyr::filter(Treatment == "0 mM") %>%
  ggplot(aes(x = BL1.H, y = BL3.H)) +
  geom_point(aes(color = interaction(Dye, Dilution))) +
  facet_wrap(~Treatment)
  #stat_summary(fun.data = "mean_cl_boot", geom = "line", size = 0.2, position = position_nudge(x = 0.1), aes(group = interaction(Dye, Dilution), color = interaction(Dye, Dilution))) +
  labs(x = "Hydrogen Peroxide Treatment", y = "Median Red Fluorescence") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        #axis.title.x = element_blank(), 
        strip.background = element_blank(),
        axis.text.x = element_text(size = 16, face = "bold"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        )
```

```{r}
p1 <- gated_data %>% 
  dplyr::filter(Dye == "SYTO9") %>% 
  dplyr::filter(Dilution == "1000 x") %>% 
   dplyr::filter(!(Treatment %in% c("0 mM")))  %>%
  mutate(Treatment = gsub(" mM", "", Treatment)) %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "skyblue2", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live Population") +
  ggtitle("SYTO9 only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot(font_size = 16) 

p2 <- gated_data %>% 
  dplyr::filter(Dye == "SYTO9") %>% 
  dplyr::filter(Dilution == "1000 x") %>% 
   dplyr::filter(!(Treatment %in% c("0 mM")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "skyblue2", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Red Negative") +
  ggtitle("SYTO9 only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot(font_size = 16) 

p3 <- gated_data %>% 
  dplyr::filter(Dye == "SYTO9") %>% 
  dplyr::filter(Dilution == "1000 x") %>% 
   dplyr::filter(!(Treatment %in% c("0 mM")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "skyblue2", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Red Negative") +
  ggtitle("SYTO9 only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot(font_size = 16) 
  
combined_plot <- plot_grid(p1, p2, p3, nrow = 1, ncol = 3, common.legend = TRUE, align = 'hv', widths = c(6, 6, 6))
print(combined_plot)
ggsave("../output/20240430-SYTO9-Cg-Log-Gated.png")
```













