---
title: "Analyze Hanxi's single stain flow results"
author: Bin He, originally by Hanxi Tang
date: "2023-10-20 (updated `r Sys.Date()`)"
output:
  html_notebook:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: hide
---

# Background
This is the flow cytometry data for SYTO9 single stained and SYTO9+PI dual stained cells. The samples are mixures of live and 1M H2O2 killed _C. glabrata_ cells. The goal of this experiment is to determine

1. Whether SYTO9 stainining is different between live and dead cells
1. Whether SYTO9 emission has significant spillover in the BL3 channel, which we use to measure PI

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
```

```{r}
old <- theme_set(theme_minimal())
```

# Import data
```{r}
data.path = "/Volumes/rdss_bhe2/User/Hanxi Tang/Flow Cytometry/101823_+-PI Filters/"
fs <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
fs0 <- fs # make a copy of the original data in case of corruption
```

Specify the flowSet phenoData
```{r}
source("../00-shared/script/20220326-simplify-names-subroutine.R")
oriNames <- sampleNames(fs)
tmp <- gsub(".fcs$", "", oriNames) %>% 
  str_split(pattern = "_| ", simplify = TRUE) 
colnames(tmp) <- c("filter", "treat") 
sample <- data.frame(filter = tmp[,4], species = tmp[, 7],
                     dead = tmp[, 8], rep = tmp[, 9]
) %>% 
  mutate(rep = ifelse(rep == "B", "B", "A"),
         dead = as.integer(gsub("%", "", dead)))
rownames(sample) <- oriNames
pData(fs) <- sample
```

# Exploratory Data Analysis
Let's take one dataset and visualize the changes in the various parameters
```{r, fig.height=6, fig.width=8}
sub <- fs[pData(fs)$species == "Cg"]
mult_format <- list(
  scale_y_continuous(name = "SSC.H (x1000)", labels = function(l) {l/1000}),
  scale_x_continuous(name = "FSC.H (x1000)", labels = function(l) {l/1000})
)
p <- ggcyto(sub, aes(x = "FSC.H", y = "SSC.H")) +
  geom_hex(aes(fill = after_stat(ncount)), bins = 80) + 
  #scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(11, "Spectral"))) +
  mult_format + facet_wrap(~paste(filter, dead, sep = "_"), scales = "fixed")
p
```

```{r}
cell.filter <- rectangleGate(filterId = "cell", 
                             "FSC.H" = c(125E3, 100E4),
                             "SSC.H" = c(0, 750E3))
# constructor for a list of filters
#cell.gates <- sapply(sampleNames(fs), function(sn)cell.filter)
p1 <- ggcyto(sub, aes(x = "FSC.H", y = "SSC.H")) + geom_hex(bins = 80) + 
  geom_gate(cell.filter) +
  facet_wrap(~H2O2, labeller = "label_both") + theme_minimal()
p1 + scale_y_continuous(name = "SSC.H (x1000)", labels = function(l) {l/1000}) + 
  scale_x_continuous(name = "FSC.H (x1000)", labels = function(l) {l/1000})
```

Remove the outliers
```{r}
sub1 <- Subset(sub, cell.filter)
p1 <- ggcyto(sub1, aes(x = "FSC.H", y = "SSC.H")) + geom_hex(bins = 80) + 
  facet_wrap(~H2O2, labeller = "label_both") + theme_minimal()
p1 + scale_y_continuous(name = "SSC.H (x1000)", labels = function(l) {l/1000}) + 
  scale_x_continuous(name = "FSC.H (x1000)", labels = function(l) {l/1000})
```
> Two populations are visible in the FSC:SSC plot. The minor population appears
> to have smaller SSC.

Try Identifying the two clusters
```{r}
tmp.dat <- sub1[[which(pData(sub)$H2O2 == 2)]]
tmp.res <- flowClust(
  tmp.dat,
  varNames = c("FSC.H", "SSC.H"),
  K = 2,
  B = 500
  )
```

```{r}
summary(tmp.res)
def.par <- par(no.readonly = TRUE) # save defaults, for resetting...
layout(matrix(c(1,2), ncol = 2))   # divide the figure into two columns
plot(tmp.res, data = tmp.dat, level = 0.8, z.cutoff = 0)
plot(density(tmp.res, data = sub[[4]]), type = "image")
par(def.par)
```

Let's split the two populations and compare their fluorescence levels
```{r}
# creates a filter object to store all settings, but doesn't perform the clustering
tmp.filter <- tmixFilter(filterId = "fsc-ssc", parameters = c("FSC.H", "SSC.H"), K = 2, B = 500)
# implement the actual clustering 
tmp.res2 <- filter(tmp.dat, tmp.filter)
```

```{r}
tmp.split <- split(tmp.dat, tmp.res2, population = list(ssc.h = 1, ssc.l = 2)) %>% as("flowSet")
# get the MFI for each subset
tmp.median <- fsApply(tmp.split, each_col, median)
tmp.median
ggcyto(tmp.split, aes(x = "FSC.H", y = "SSC.H")) + geom_hex(bins = 80) + theme_minimal()
ggcyto(tmp.split, aes(x = "BL1.H", y = "BL3.H")) + geom_hex(bins = 80) + scale_x_logicle() + scale_y_logicle() +
  theme_minimal()
```
> Interestingly, the SSC High population appear to be live and SSC low population are mostly dead
> Also of interest is the presence of a subpopulation with very low staining by FungaLight

To examine the low-staining pop further, I will make a gate on BL1.H and BL3.H
```{r}
unstained.filter <- rectangleGate(filterId = "unstained", 
                                  list("BL1.H" = c(0, 100),
                                       "BL3.H" = c(0, 100)))
unstained.res <- filter(tmp.dat, unstained.filter)
unstained.pop <- split(tmp.dat, unstained.res) %>% as("flowSet")
ggcyto(unstained.pop, aes(x = "FSC.H")) + geom_density(aes(color = factor(name)), linewidth = 1.5) + 
  scale_color_manual("Unstained", values = c("unstained-" = "gray20", "unstained+" = "red")) +
  facet_wrap(~NULL) + theme_minimal(base_size = 16)
```
> The unstained population of events seem to be smaller in size compared with the stained population
> wonder if this has to do with a transient population of cells with physiological characteristics
> that make them both smaller and less permeable to both dyes.
> do these cells represent dead cells or live cells, or an equal mixture of both?

```{r, fig.height=6, fig.width=8}
p2 <- ggcyto(sub1, aes(x = BL1.H, y = BL3.H)) + 
  geom_hex(aes(fill = after_stat(ncount)), bins = 80) + 
  facet_wrap(~H2O2, labeller = "label_both") + theme_minimal()
p2 + scale_y_logicle() + scale_x_logicle()
```
> overall, we see that the two extremes are well-characterized, i.e., mock treated sample has a very minor 
> population of dead cells, which can be recognized by looking at the 1M treated sample.
> the 

Extract MFI for BL1.H across all flowFrames
```{r}
tmp <- fsApply(fs, each_col, median, use.exprs = TRUE)
```

# Gatting
```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(fs, aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2)# + ggcyto_par_set(limits = "instrument")
```
> Most events in all samples fell within this outlier/instrument gate.

Add this gate to the GatingSet

```{r}
gs <- GatingSet(fs) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

Let's examine how this gate intersected with the FSC.H vs FSC.W plot (for singlets)

```{r}
p1 <- ggcyto(gs[[1]], aes(x = FSC.H, y = FSC.W), subset = "root") + geom_hex(bins = 128)
p2 <- ggcyto(gs[[1]], aes(x = FSC.H, y = FSC.W), subset = "-outlier") + geom_hex(bins = 128)
plot_grid(as.ggplot(p1), as.ggplot(p2), ncol = 2)

```
Gate for singlets
```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(fs[[1]], outlier.gate)
#singlet.gate <- gate_flowclust_2d(ex, "FSC.H", "FSC.W", filterId = "singlet", K = 2, quantile = 0.8)
# switch to a polygongate
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 60, 75, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

Visualize the data
```{r}
ggcyto(gs[sample(shortNames,8)], aes(x = BL1.H, y = BL3.H), subset = "-outlier") + geom_hex(bins = 64) +
  facet_wrap(~name, ncol = 4) + scale_x_logicle() + scale_y_logicle() + theme_bw()
```
# Gate for Dead population
> Dead population is defined based on 1M treated sample.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(fs[[15]], singlet.gate)
polygon <- matrix(c(0, 0, 10^3.1, 10^3.1, 10^2.9, 10^2.4, 10^1.8,
                    10^3.2, 10^3.5, 1e4, 10^3.4, 10^3, 10^2.6, 10^2.2), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
dead.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(dead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, dead.gate, parent = "singlet", name = "dead")
recompute(gs)
```

# Gate for Live population
> live population is gated based on the mock treated sample

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(fs[[12]], singlet.gate)
polygon1 <- matrix(c(1e2, 1e2, 10^2.5, 10^4, 1e3, 
                    0, 1e2, 10^2.1, 10^3, 0), ncol = 2)
colnames(polygon1) <- c("BL1.H", "BL3.H")
live.gate <- polygonGate(filterId = "live", boundaries = polygon1)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(live.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, live.gate, parent = "singlet", name = "live")
recompute(gs)
```

# Gate for Unstained population
> Unstained population is defined as population below 10^2 in both channels (in the noise range).

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(fs[[7]], singlet.gate)
polygon2 <- matrix(c(0, 1e2, 1e2, 0,
                    0, 0, 1e2, 1e2), ncol = 2)
colnames(polygon2) <- c("BL1.H", "BL3.H")
unstained.gate <- polygonGate(filterId = "unstained", boundaries = polygon2)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(unstained.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, unstained.gate, parent = "singlet", name = "unstained")
recompute(gs)
```

# Gate for stained population
> stained population is defined based populatio  above 10^2 in both channels.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(fs[[7]], singlet.gate)
polygon3 <- matrix(c(1e2, 1e6, 1e6, 1e2,
                    0, 0, 1e6, 1e6), ncol = 2)
colnames(polygon3) <- c("BL1.H", "BL3.H")
stained.gate <- polygonGate(filterId = "stained", boundaries = polygon3)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(stained.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, stained.gate, parent = "singlet", name = "stained")
recompute(gs)
```


# Gate for red positive population
> Red population is defined as red positive events in the stained population. It's gated based on the 1M treated population

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(fs[[16]], stained.gate)
polygon4 <- matrix(c(1e2, 1e6, 1e6, 1e2,
                    10^2.5, 10^2.5, 1e6, 1e6), ncol = 2)
colnames(polygon4) <- c("BL1.H", "BL3.H")
RP.gate <- polygonGate(filterId = "RP", boundaries = polygon4)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(RP.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, RP.gate, parent = "stained", name = "Red_Positive")
recompute(gs)
```


# Extract event counts, write into tibble, export into csv

```{r}
smaller <- Subset(fs, live.gate)
summary(smaller)
```



stats <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)%>% 
 mutate(Live_green_median = live_median)
print(stats)
saving_path <- file.path("/space/htang5/Documents/092223_Cg FGL Analysis", "cg_flow_median.csv")
write_csv(stats, saving_path)
```


> CFU data is manually attached to the csv, and the csv is filtered and re-imported as a csv below

# Re-import data, add variables, plot

```{r}
mydata <- read.table("/space/htang5/Documents/092223_Cg FGL Analysis/Cg_FGL-CFU 09 - cg_flow.tsv", sep="\t", header=TRUE)
mydata$CFU_Survival <- mydata$CFU_Live / mydata$CFU_Total
mydata$Percent_Live <- mydata$n_live / (mydata$n_singlet - mydata$n_unstained)
mydata$Percent_Dead <- mydata$n_dead / mydata$n_singlet
mydata$Percent_Unstained <- mydata$n_unstained / mydata$n_singlet
mydata$Percent_RedPos_Stained <- mydata$n_Red_Positive / mydata$n_stained
mydata$Percent_Red_neg <- 1- (mydata$n_Red_Positive / mydata$n_stained)

mydata %>%
  filter(Treatment %in% c("20 mM", "40 mM", "60 mM", "80 mM", "100 mM"))  %>% 
  ggplot(aes(x = Percent_Live, y = CFU_Survival, color = Treatment)) +
  geom_point()
ggplot(mydata, aes(x = Percent_Unstained, y = CFU_Survival)) +  geom_point ()
ggplot(mydata, aes(x = Percent_Dead, y = CFU_Survival)) +  geom_point ()
ggplot(mydata, aes(x = Percent_RedPos_Stained, y = CFU_Survival)) +  geom_point ()
ggplot(mydata, aes(x = Percent_Red_neg, y = CFU_Survival)) +  geom_point ()
ggplot(mydata, aes(x = sample, y = n_singlet)) +
  geom_bar(stat = "identity") +  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_hline(yintercept = 1000, linetype = "dashed", color = "red")
```
```{r}
mydata %>%
  mutate(Treatment = factor(Treatment, levels = c("2 mM", "4 mM", "6 mM", "8 mM", "10 mM", "20 mM", "40 mM", "60 mM", "80 mM", "100 mM"))) %>%
ggplot(aes(x = Treatment, y = Percent_Live)) + 
  geom_boxplot()+
  geom_point()+
  theme_minimal() +
  labs(x = "Treatment Hydrogen Peroxide Concentration (mM)",
       y = "Gated Live Population Percentage") 
```


# Calculate R^2 for each correlation
```{r}
sample <- c(sample)

lm.res <- mydata %>%
  as.data.frame() %>%
  lm(CFU_Survival ~ Percent_Live, data = .)  

summary(lm.res)

lm.res2 <- mydata %>%
  as.data.frame() %>%
  lm(CFU_Survival ~ Percent_Dead, data = .)  

summary(lm.res2)

lm.res3 <- mydata %>%
  as.data.frame() %>%
  lm(CFU_Survival ~ Percent_RedPos_Stained, data = .)  

summary(lm.res3)

lm.res4 <- mydata %>%
  as.data.frame() %>%
  lm(CFU_Survival ~ Percent_Red_neg, data = .)  

summary(lm.res4)
```


```{r}
library(ggplot2)
library(viridis)
ggplot(mydata, aes(x = Percent_Live, y = CFU_Survival)) + 
  geom_point(aes(color = Treatment), size = 3) +
  stat_smooth(method = "lm") +
  labs(x = "Gated Live Population Percentage",
       y = "CFU Survival")  +
  theme_minimal() +
  theme(axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        axis.line = element_line(color = "black"),
        legend.text = element_text(size = 14, face = "bold"),   # Set legend label size and boldness
        legend.title = element_text(size = 14, face = "bold"))  # Set legend title size and boldness

```

```{r}
ggplot(mydata_filtered, aes(x = Percent_Dead, y = CFU_Survival)) + 
  geom_point()+ scale_y_sqrt()
```














