---
title: "Fig 3 Graphs"
author: Bin He, originally by Hanxi Tang
date: "2023-09-30 (updated `r Sys.Date()`)"
output:
  html_notebook:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: hide
---

# Background
This is the flow cytometry and CFU data for Cg, Sc, Kl. Cg, Sc, Kl treated with 0-1M H2O2 are stained with Fungalight 1, run through flow and plated for CFU (details in HT's ELN). Goal is to correlate Flow vs CFU and generate graphs for figure 3.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
```

# Import data
```{r}
data.path = "/Volumes/rdss_bhe2/User/Hanxi Tang/Flow Cytometry/092923-FungaLight-Cgla-analysis/"
fs <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
fs0 <- fs # make a copy of the original data in case of corruption
```

Specify the flowSet phenoData
```{r}
source("../00-shared/script/20220326-simplify-names-subroutine.R")
oriNames <- sampleNames(fs)
tmp <- str_split(oriNames, pattern = "_", simplify = TRUE)[,c(1, 5)] 
colnames(tmp) <- c("date", "treat") 
sample <- data.frame(tmp) %>% 
  mutate(rep = ifelse(str_sub(treat, 1, 1) == "B", "B", "A"),
         treat = gsub(".fcs", "", treat) %>% gsub("^B ", "", .),
         H2O2 = case_match(treat,
                             "1M" ~ "1000",
                             "Mock" ~ "0",
                             .default = treat),
         H2O2 = factor(H2O2, levels = sort(unique(as.numeric(H2O2)))),
         set = factor(paste0(date, rep)),
         name = paste(date, rep, H2O2, "mM", sep = "_"))
rownames(sample) <- oriNames
pData(fs) <- sample
sets <- levels(sample$set)
```

# Exploratory Data Analysis
Let's take one dataset and visualize the changes in the various parameters

```{r}
# define the rectangular gate
cell.filter <- rectangleGate(filterId = "cell", 
                             "FSC.H" = c(125E3, 100E4),
                             "SSC.H" = c(0, 750E3))
```

Plot all sets
```{r, fig.height=6, fig.width=8}
png("img/20230930-")
for(set in sets){
  sub <- fs[pData(fs)$set == set]
  # constructor for a list of filters
  #cell.gates <- sapply(sampleNames(fs), function(sn)cell.filter)
  p1 <- ggcyto(sub, aes(x = "FSC.H", y = "SSC.H")) + geom_hex(bins = 80) + 
    geom_gate(cell.filter) +
    facet_wrap(~H2O2, labeller = "label_both") + theme_minimal() +
    labs(title = set) +
    scale_y_continuous(name = "SSC.H (x1000)", labels = function(l) {l/1000}) + 
    scale_x_continuous(name = "FSC.H (x1000)", labels = function(l) {l/1000})
  print(p1)
}
```

> Two populations are visible in the FSC:SSC plot. The minor population with smaller SSC values are more prominent in the more severely stressed cells.

Try Identifying the two clusters
```{r}
tmp.dat <- fs[pData(fs)$set == "091523A" & pData(fs)$H2O2 == 8][[1]]
tmp.res <- flowClust(
  tmp.dat,
  varNames = c("FSC.H", "SSC.H"),
  K = 2,
  B = 500
  )
```

```{r}
summary(tmp.res)
def.par <- par(no.readonly = TRUE) # save defaults, for resetting...
layout(matrix(c(1,2), ncol = 2))   # divide the figure into two columns
plot(tmp.res, data = tmp.dat, level = 0.8, z.cutoff = 0)
plot(density(tmp.res, data = sub[[4]]), type = "image")
par(def.par)
```

Let's split the two populations and compare their fluorescence levels
```{r}
# creates a filter object to store all settings, but doesn't perform the clustering
tmp.filter <- tmixFilter(filterId = "fsc-ssc", parameters = c("FSC.H", "SSC.H"), K = 2, B = 500)
# implement the actual clustering 
tmp.res2 <- filter(tmp.dat, tmp.filter)
```

```{r}
tmp.split <- split(tmp.dat, tmp.res2, population = list(ssc.h = 1, ssc.l = 2)) %>% as("flowSet")
# get the MFI for each subset
tmp.median <- fsApply(tmp.split, each_col, median)
tmp.median
ggcyto(tmp.split, aes(x = "FSC.H", y = "SSC.H")) + geom_hex(bins = 80) + theme_minimal()
ggcyto(tmp.split, aes(x = "BL1.H", y = "BL3.H")) + geom_hex(bins = 80) + scale_x_logicle() + scale_y_logicle() +
  theme_minimal()
```
> Interestingly, the SSC High population appear to be live and SSC low population are mostly dead
> Also of interest is the presence of a subpopulation with very low staining by FungaLight. This population is mostly in the SSC-high, thus live portion of the sample.

To examine the low-staining pop further, I will make a gate on BL1.H and BL3.H
```{r}
unstained.filter <- rectangleGate(filterId = "unstained", 
                                  list("BL1.H" = c(0, 100),
                                       "BL3.H" = c(0, 100)))
unstained.res <- filter(tmp.dat, unstained.filter)
unstained.pop <- split(tmp.dat, unstained.res) %>% as("flowSet")
ggcyto(unstained.pop, aes(x = "FSC.H")) + geom_density(aes(color = factor(name)), linewidth = 1.5) + 
  scale_color_manual("Unstained", values = c("unstained-" = "gray20", "unstained+" = "red")) +
  facet_wrap(~NULL) + theme_minimal(base_size = 16)
```
> The unstained population of events seem to be a bit smaller in size compared with the stained population.
> wonder if this has to do with a transient population of cells with physiological characteristics
> that make them both smaller and less permeable to both dyes.

```{r, fig.height=6, fig.width=8}
for(set in sets){
  sub <- fs[pData(fs)$set == set]
  # constructor for a list of filters
  #cell.gates <- sapply(sampleNames(fs), function(sn)cell.filter)
  p2 <- ggcyto(sub, aes(x = BL1.H, y = BL3.H)) + 
    geom_hex(aes(fill = after_stat(ncount)), bins = 80) + 
    facet_wrap(~H2O2, labeller = "label_both") +
    scale_y_logicle() + scale_x_logicle() +
    labs(title = set) +
    theme_minimal()
  print(p2)
}
```
> We can observe some trends with increasing H2O2 severity, such as an increase in the 
> dead population (top left), more cells exhibiting a higher green and red within the "live"
> population, possibly due to increased permeability to SYTO9, which is much more fluorescent
> than PI. Eventually, the population does migrate to the dead gate. Despite these useful trends
> however, there is a lot of run-to-run variability, to the point where confident assessment of
> the patterns is not possible.

