---
title: "PI-Concentration-Distinguishing-Power"
output: html_document
date: "2023-12-18"
---

# Background
This is the flow cytometry data for 0-1000mM treated Cg stained with various concentrations of PI+SYTO9, and run through flow cytometry (details in ELN). The goal is to establish PI+SYTO's distinguishing power for live vs dead cells.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
require(ggrdiges)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "/space/htang5/Documents/20231211_Cg Flow + CFU/Raw"
dat.f2b <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
require(PTXQC)
oriNames <- sampleNames(dat.f2b)
tmpb <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 6)] 
colnames(tmpb) <- c("Date", "Treatment") 
sampleb <- data.frame(tmpb) %>% 
  mutate(Treatment = factor(gsub(".fcs", "", Treatment), levels = as.character(c(0, 20, 40, 60, 80, 100, 1000))) %>% paste(" mM", sep = "")) 
rownames(sampleb) <- oriNames
pData(dat.f2b) <- sample
print(pData(dat.f2b))
write_tsv(pData(dat.f2b), file = "../Input/20240104-fig-2b-PI+SYTO9-raw-data-ht.tsv")
```

# Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier 

```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat.f2b [1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat.f2b) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

>Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat.f2b[[1]], outlier.gate)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```
> Gate for Dead population
Dead population is defined by having red fluorescence value above 10^2.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f2b[[10]], singlet.gate)
polygon <- matrix(c(0, 10^3.6, 10^3.6, 0,
                    10^2.15, 10^2.15, 1e6, 1e6), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
dead.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(dead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r}
gs_pop_add(gs, dead.gate, parent = "singlet", name = "dead")
recompute(gs)
```

> Gate for Live
live population is gated based on the mock treated sample

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f2b[[8]], singlet.gate)
polygon1 <- matrix(c(0, 10^2.8, 10^2.8, 0,
                    0, 0, 10^2.15, 10^2.15), ncol = 2)
colnames(polygon1) <- c("BL1.H", "BL3.H")
live.gate <- polygonGate(filterId = "live", boundaries = polygon1)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(live.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, live.gate, parent = "singlet", name = "live")
recompute(gs)
```

>Gate for Intermediate
live intermediate is gated based on 100mM treated sample

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f2b[[7]], singlet.gate)
polygon2 <- matrix(c(10^2.8, 10^4, 10^4, 10^2.8,
                    0, 0, 10^3, 10^3), ncol = 2)
colnames(polygon2) <- c("BL1.H", "BL3.H")
inter.gate <- polygonGate(filterId = "Inter", boundaries = polygon2)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(inter.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```

Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, inter.gate, parent = "singlet", name = "Intermediate")
recompute(gs)
```

>Exported gated event counts

```{r}
gated_stats_f2b <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)

gated_stats_f2b$percent_dead <- gated_stats_f2b$n_dead / gated_stats_f2b$n_singlet
gated_stats_f2b$percent_live <- gated_stats_f2b$n_live / gated_stats_f2b$n_singlet
gated_stats_f2b$percent_inter <- gated_stats_f2b$n_Intermediate / gated_stats_f2b$n_singlet
gated_stats_f2b$FGS <- gated_stats_f2b$percent_live + 0.5*gated_stats_f2b$percent_inter - 1*gated_stats_f2b$percent_dead

gated_data <- cbind(pData(dat.f2b), gated_stats_f2b) %>% 
    as.tibble() %>% relocate(name, .after = last_col())

write_tsv(gated_data, file = "../Input/20240104-fig-2b-PI+SYTO9-Gated-ht.tsv")
print(gated_data)
```


# Output Graph Production
>Several output graphs are produced, and are stored in the Output folder.

> Plot FGS vs [H2O2]

```{r}
pf2b <- gated_data %>%
  #dplyr::filter(Dye == "PI") %>% 
  #dplyr::filter(Treatment == "0 mM") %>%
  mutate(Treatment = factor(Treatment, levels = c("0", "20", "40", "60", "80", "100", "1000"))) %>%
  ggplot(aes(x = Treatment, y = percent_inter)) +
  geom_boxplot() +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.2)) +
  theme_minimal()
pf2b +
  labs(x = "Hydrogen Peroxide Treatment Concentration (mM)", y = "FungaLight Score") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 12, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 16, face = "bold", color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 16, face = "bold")) 
#ggsave("../output/20240104-PI+SYTO9_Gated_Plot.png", width = 7, height = 5)
```















