---
title: "PI-Concentration-Distinguishing-Power"
output: html_document
date: "2023-12-18"
---

# Background
This is the flow cytometry data for 0-1000mM treated Cg stained with various concentrations of PI alone, and run through flow cytometry (details in ELN). The goal is to establish PI's distinguishing power for live vs dead cells.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
require(ggrdiges)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "/space/htang5/Documents/112723_SP Dilutions/Raw"
dat.f2 <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
require(PTXQC)
oriNames <- sampleNames(dat.f2)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 5, 6, 7)] 
colnames(tmp) <- c("Date", "Treatment", "Dilution", "Dye") 
sample <- data.frame(tmp) %>% 
 mutate(Dye = ifelse(Dye == "p.fcs", "PI", "SYTO9"),
    Dilution = factor(gsub(".fcs", "", Dilution), levels = as.character(c(250, 500, 1000, 2000))) %>% paste(" x", sep = ""),
    Treatment = factor(Treatment, levels = as.character(c(0, 10, 100, 1000))) %>% paste(" mM", sep = ""))
rownames(sample) <- oriNames
pData(dat.f2) <- sample
print(pData(dat.f2))
write_tsv(pData(dat.f2), file = "../Input/20240104-fig-2-PI-raw-data-ht.tsv")
```

# Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier 

```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat.f2 [1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat.f2) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

>Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat.f2[[1]], outlier.gate)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

>Gate for Red Positive
RP population is defined by having red fluorescence value above 10^2.15.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f2[[25]], singlet.gate)
polygon <- matrix(c(0, 10^3.5, 10^3.5, 0,
                    10^2.15, 10^2.15, 1e6, 1e6), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
RP.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(RP.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, RP.gate, parent = "singlet", name = "RedPositive")
recompute(gs)
```

>Exported gated event counts

```{r}
gated_stats_f2 <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)
gated_stats_f2$percent_RP <- stats$n_RedPositive / stats$n_singlet
gated_data <- cbind(pData(dat.f2), gated_stats_f2) %>% 
    as.tibble() %>% relocate(name, .after = last_col())

write_tsv(gated_data, file = "../Input/20240104-fig-2-PI-Gated-ht.tsv")
print(gated_data)
```


# Output Graph Production
>Several output graphs are produced, and are stored in the Output folder.

>Visualizing flow raw data (not included in output)

```{r}
sub <- dat.f2[pData(dat.f2)$Dye == "PI" & pData(dat.f2)$Date == "112023"] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  facet_grid(Dilution ~ Treatment, 
             labeller = as_labeller(c("0 mM" = "0 mM", "10 mM" = "10 mM", "100 mM" = "100 mM", "1000 mM" = "1000 mM", 
                                      "250 x" = "250x", "500 x" = "500x", "1000 x" = "1000x", "2000 x" = "2000x"))) + 
  theme_minimal()
p2 + scale_x_logicle() + scale_y_logicle() +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence") +
  panel_border(color = "black", size = 1) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        ) + 
  theme(strip.text = element_text(size = 12, face = "bold")) 
#ggsave("../output/20231228-PI-Dilutions-Raw-Flow-Plot.png", width = 5, height = 3)
```


> Ridges plot for demsontrationg gating scheme

```{r}
# Subset and filter data
sub <- dat.f2[pData(dat.f2)$Dye == "PI" & pData(dat.f2)$Date == "112023" & pData(dat.f2)$Dilution == "1000 x"] 
sg <- Subset(sub, singlet.gate)

# Create the base plot components
p0 <- list(
  geom_density_ridges(aes(y = Treatment)), 
  scale_x_flowjo_biexp(),
  facet_null(),
  theme_bw(base_size = 12)
)

p01 <- list(
  xlab("FungaLight Red Fluorescence"),
  ylab("Treatment Hydrogen Peroxide Concentration"),
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 14, color = "black", face = "bold"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 14, color = "black", face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold")
        )
)

# Create individual ggcyto plots
p4 <- ggcyto(sg, aes(x = BL3.H)) + p0 + p01
print(p4)
ggsave("../output/20240104-PI-Distinguishing-Raw-Ridges-Plot.png", width = 5, height = 5)
```
> Plot % Red Positive vs [H2O2]

```{r}
pf1a <- gated_data %>%
  dplyr::filter(Dye == "PI") %>% 
  #dplyr::filter(Treatment == "0 mM") %>%
  mutate(Dilution = factor(Dilution, levels = c("250 x", "500 x", "1000 x", "2000 x"))) %>%
  ggplot(aes(x = Treatment, y = percent_RP)) +
  facet_wrap(~ Dilution, scale = "free_y") +
  geom_boxplot() +
  geom_point(stroke = 1, size = 2, position = position_jitter(0.2)) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
pf1a +
  labs(x = "Hydrogen Peroxide Treatment Concentration", y = "Gated Percent Red Positive Cells") +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 12, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 16, face = "bold", color = "black"),
        axis.title.x = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 16, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 16, face = "bold")) 
ggsave("../output/20240104-PI-Dilutions-Gated_RP-Plot.png", width = 7, height = 5)
```















