---
title: "PI vs PI+SYTO9 Log Correlation"
output: html_document
date: "2023-12-18"
---

# Background
This is the flow cytometry data for post treatment phase Cg stained with PI/PI+SYTO9, and run through flow cytometry (details in ELN), and plated for CFU. The goal is to establish FungaLight's distinguishing power and correlate with plating based method.

```{r setup, message=FALSE}
require(tidyverse)
require(flowCore)
require(flowClust)
require(openCyto)
require(ggcyto)
require(cowplot)
require(ggrdiges)
```

# Import data
> FCS files are stored in RDSS/user/flow cytometry, FCS is read and write into an input tsv table. The tsv file is avaliable from the Input folder.

```{r}
# use relative path to make it easier for collaboration
data.path = "/space/htang5/Documents/20240225-28"
dat.f1g <- read.flowSet(path = data.path, transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

Simplify the sample names

```{r}
require(PTXQC)
oriNames <- sampleNames(dat.f1g)
tmp <- str_split(oriNames, pattern = "[ _]+", simplify = TRUE)[,c(1, 7, 8)]
colnames(tmp) <- c("Date", "Treatment", "Dye") 
sample <- data.frame(tmp) %>% 
  mutate(
    Dye = ifelse(Dye == "p.fcs", "PI", ifelse(Dye == "b.fcs", "Both", Dye)), 
    Treatment = factor(Treatment, levels = c(0, "5C", 0.6, 6, 1000)))
rownames(sample) <- oriNames
pData(dat.f1g) <- sample
print(pData(dat.f1g))
#write_tsv(pData(dat.f1g), file = "/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/02-FungaLight-Distinguishing-Power/Input/20240303-Sc-PI-Both-Log-Correlation-raw-data-ht.tsv")
```


# Gatting for Singlets
>The following gaphing steps are used to gate singlets by FSC and SSC values. Only singlets are included in analysis.

>Gate for outlier 

```{r}
outlier.gate <- rectangleGate(filterId = "-outlier", "FSC.H" = c(1.2e5, 1e6), "SSC.H" = c(1e2, 1e6))
ggcyto(dat.f1g [1], aes(x = FSC.H, y = SSC.H), subset = "root") +
  geom_hex(bins = 64) + geom_gate(outlier.gate) + facet_wrap(~name, ncol = 2) + ggcyto_par_set(limits = "instrument")
```
Add gate to GS
```{r}
gs <- GatingSet(dat.f1g) # create a GatingSet
gs_pop_add(gs, outlier.gate, parent = "root")
recompute(gs)
```

>Gate for singlets

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(30,300)))
ex <- Subset(dat.f1g[[1]], outlier.gate)
polygon <- matrix(c(1e5, 1e5, 1e6, 1e6, 
                    60, 105, 135,60), ncol = 2)
colnames(polygon) <- c("FSC.H", "FSC.W")
singlet.gate <- polygonGate(filterId = "singlet", boundaries = polygon)
ggcyto(ex, aes(x = FSC.H, y = FSC.W)) + geom_hex(bins = 128) + geom_gate(singlet.gate) + geom_stats() + scPars
```


Add this gate to the gatingSet
```{r}
gs_pop_add(gs, singlet.gate, parent = "-outlier", name = "singlet")
recompute(gs)
```

> Gate for Red Negative Population
This is the live population with red fluorescence below 10^2.2

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g, singlet.gate)
polygon <- matrix(c(0, 10^5, 10^5, 0,
                    10^2.2, 10^2.2, 0, 0), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
live.gate <- polygonGate(filterId = "live", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(live.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() + facet_grid (Treatment ~ Dye)
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, live.gate, parent = "singlet", name = "live")
recompute(gs)
```

> Gate for Red Positive Intermediates
This is the intermediate population on the right side

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g[[10]], singlet.gate)
polygon <- matrix(c(10^5, 10^3,10^5,
                    10^2.2, 10^2.2, 10^4), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
inter.gate <- polygonGate(filterId = "inter", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(inter.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, inter.gate, parent = "singlet", name = "inter")
recompute(gs)
```

> Gate for Red Positive Dead
This is the dead population on the right side

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g[[20]], singlet.gate)
polygon <- matrix(c(10^3, 10^2.5,10^3, 10^5,
                    10^4.5, 10^2.2, 10^2.2, 10^4), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
dead.gate <- polygonGate(filterId = "dead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(dead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, dead.gate, parent = "singlet", name = "dead")
recompute(gs)
```

> Gate for Dead Population - PI only
RP population is defined by having red fluorescence value above 10^2.15.

```{r fig.width=4, fig.height=4}
scPars <- ggcyto_par_set(limits = list(x = c(0,1e6), y = c(0,1e6)))
ex <- Subset(dat.f1g[[14]], singlet.gate)
polygon <- matrix(c(0, 10^3.6, 10^3.6, 0,
                    10^2.15, 10^2.15, 1e6, 1e6), ncol = 2)
colnames(polygon) <- c("BL1.H", "BL3.H")
PIdead.gate <- polygonGate(filterId = "PIdead", boundaries = polygon)
ggcyto(ex, aes(x = BL1.H, y = BL3.H)) + geom_hex(bins = 128) + geom_gate(PIdead.gate) + geom_stats() + scPars + scale_x_logicle() + scale_y_logicle() 
```
Add this gate to the gatingSet

```{r eval=FALSE, include=FALSE}
gs_pop_add(gs, PIdead.gate, parent = "singlet", name = "PIDead")
recompute(gs)
```

>Exported gated event counts

```{r}
gated_stats_f1g <- gs_pop_get_stats(gs) %>% 
  as_tibble() %>% 
  mutate(pop = gsub(".*/", "", pop), pop = gsub("-outlier", "cells", pop)) %>% 
  pivot_wider(names_from = pop, names_prefix = "n_", values_from = count)
gated_stats_f1g$percent_Dead <- gated_stats_f1g$n_dead / gated_stats_f1g$n_singlet
gated_stats_f1g$percent_Live <- gated_stats_f1g$n_live / gated_stats_f1g$n_singlet
gated_stats_f1g$percent_inter <- gated_stats_f1g$n_inter / gated_stats_f1g$n_singlet
gated_stats_f1g$FGS <- 1*gated_stats_f1g$percent_Live + 0.5*gated_stats_f1g$percent_inter - gated_stats_f1g$percent_Dead
gated_data <- cbind(pData(dat.f1g), gated_stats_f1g) %>% 
    as.tibble() %>% relocate(name, .after = last_col())
write_tsv(gated_data, file = "../Input/20240303-Sc-PI-Both-Distinguishing-Power-gated-ht.tsv")
print(gated_data)
```
# Extract MFI of intermediate population green channel
>MFI are extracted and wrote into tsv. The tsv is stored in the input folder.

```{r}
interpop <- Subset(dat.f1g, inter.gate)
tmp <- fsApply(interpop, each_col, median)
combined_data <- cbind(gated_data, tmp) %>%
    as.tibble() %>% relocate(name, .after = last_col())

#write_tsv(combined_data, file = "../Input/20240124_four_variable_matrix.tsv")
print(combined_data)
```


# Output Graph Production
>Several output graphs are produced, and are stored in the Output folder.

```{r fig.width=8, fig.height=8}
sub <- dat.f1g[pData(dat.f1g)$Dye == "Both" & 
                 pData(dat.f1g)$Date == "022624" & 
                 (pData(dat.f1g)$Treatment == "0" |
                 pData(dat.f1g)$Treatment == "0.6" |
                  pData(dat.f1g)$Treatment == "6" |
                 pData(dat.f1g)$Treatment == "1000")] 
sg <- Subset(sub, singlet.gate)
p2 <- sg %>% 
  ggcyto(aes(x = "BL1.H", y = "BL3.H")) + 
  geom_hex(bins = 80, aes(fill = ..ncount..)) + 
  facet_wrap(~ Treatment, nrow = 2, labeller = as_labeller(c("0" = "0 mM", "0.6" = "0.6 mM", "6" = "6 mM", "1000" = "1000 mM"))) + 
  theme_minimal()
p2 + scale_x_logicle(breaks = c(100, 1000, 10000, 100000)) + scale_y_logicle(breaks = c(100, 1000, 10000, 10000, 100000)) +
  xlab("FungaLight Green Fluorescence") + ylab("FungaLight Red Fluorescence")  +
  panel_border(color = "black", size = 2) +
  theme(axis.line = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 25, face = "bold", color = "black"),  # Adjust size and face (bold)
        axis.text.y = element_text(size = 25, face = "bold", color = "black"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold")
        )+ 
  theme(strip.text = element_text(size = 25, face = "bold")) 
ggsave("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/02-FungaLight-Distinguishing-Power/output/20240430-Sc-log-Raw-Flow-Plot.png")
```

```{r}
combined_data <- read_tsv("/space/htang5/Documents/FUN1_data_analysis/biol-4386-course-project-htang5/Methods Paper Writing/02-FungaLight-Distinguishing-Power/Input/20240303-Sc-PI-Both-Distinguishing-Power-CFU-ht.tsv") %>% 
mutate(Treatment = factor(Treatment, levels = c(0, "5C", 0.6, 6, 1000)))

p1 <- combined_data %>% 
  dplyr::filter(Dye == "PI") %>% 
   dplyr::filter(!(Treatment %in% c("5C", "0")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightpink", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Red Negative") +
  ggtitle("PI only") +  
  theme(plot.title = element_text(hjust = 30)) +
  theme_cowplot(font_size = 16) 
  

p2 <- combined_data %>% 
  dplyr::filter(Dye == "Both") %>% 
   dplyr::filter(!(Treatment %in% c("5C", "0")))  %>%
  ggplot(aes(x = Treatment, y = CFU_Survival))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightpink", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
   scale_y_continuous(labels = scales::percent_format()) +
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Viable") +
  ggtitle("CFU") +  
  theme_cowplot(font_size = 16) 
  theme(plot.title = element_text(hjust = 15)) 

p3 <- combined_data %>% 
  dplyr::filter(Dye == "Both") %>% 
   dplyr::filter(!(Treatment %in% c("5C", "0")))  %>%
  ggplot(aes(x = Treatment, y = percent_Live))  + 
  geom_bar(stat = "summary", fun = "mean", fill = "lightpink", alpha = 0.7) +
  geom_point(size = 3, position = position_jitter(0.2)) + 
  labs(x = bquote(H[2]*O[2]~(mM)), y = "% Live Population") +
  ggtitle("FungaLight") +  
  theme(plot.title = element_text(hjust = 15)) +
  theme_cowplot(font_size = 16) 

combined_plot <- plot_grid(p1, p2, p3, nrow = 1, ncol = 3, common.legend = TRUE, align = 'hv')

print(combined_plot)
ggsave("../output/20240427-Sc-Log-Gated.png")
```





