---
title: "Make CFU + FUN1 Plot"
output: html_document
date: "2023-04-08"
---
# Background
This is the flow cytometry data from 02/24/23. Data includes Cg treated with various [H2O2], and corresponding CFU data. The goal is to generate a graph with mean fluorescence on x-axis and CFU survival on y-axis.

### Import data
```{r}
## Load libraries
library(tidyverse)
library(flowCore)


data.path = "C:/Users/tangh/Downloads/0224RG"
fs <- read.flowSet(path = data.path, pattern = "*.fcs",
                   transformation = FALSE,  # the original values are already linearized. 
                   emptyValue = FALSE,  alter.names = TRUE,   # change parameter names to R format
                   column.pattern = ".H|FSC|SSC") # only load the height variables for the fluorescent parameters
```

### Write Data into table
```{r}
summary_table <- as.data.frame(summary(fs))
transposed_table <- t(summary_table)
write.csv(transposed_table, "H2O2fs.csv", row.names = TRUE)
```

>Some data cleaning and CFU data attachement is done in Excel due to the fact that CFU count data is stored in the Excel format.

### Plotting
```{r}
MeanFL <- read.csv("C:/Users/tangh/Downloads/FUN1H2O2_Cleaned.csv")
MeanFL
ggplot(MeanFL, aes(x = FUN1.Green_Mean, y = CFU_Survival)) +  geom_point ()
ggplot(MeanFL, aes(x = FUN1.Red_Mean, y = CFU_Survival)) +  geom_point ()
```










